---
title: "Final Project - Markov Chain Analysis with R"
author: "G.M. Arafat Rahman"
date: '2022-08-18'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Markov Chain Analysis with Weather Dataset

> **Historical Daily Weather Data 2020**
>
> This dataset contains historical daily weather data for 163 countries(with provincial data for some) from Jan 1, 2020 up to April 21, 2020. The countries and provinces were chosen based on the Johns Hopkins COVID-19 dataset. It contains various features such as temperature, pressure, humidity, ozone levels, visibility, precipitation, etc. This dataset was collected from Kaggle.

## Calling necessary libraries

> I have used the markovchain package for demonstrating markov chain.

```{r}
library(dplyr)
library(data.table)
library(markovchain)
library(diagram)
library(expm)
library(pracma)

```

> Now loading the data-set.

```{r}
df_weather = read.csv("C:\\Users\\USER\\Downloads\\STA301\\daily_weather_2020.csv")
```

> We can see the countries those are included in the data-set.

```{r}
unique(df_weather$Country.Region)
```

> I took Bangladesh's weather data to apply Markov chain.

```{r}

df_Bangladesh = filter(df_weather,Country.Region=="Bangladesh")
df_Bangladesh = select(df_Bangladesh, Country.Region, time, summary , icon)
```

> We can see the unique weather condition in Bangladesh so that we can build the states for markov chain.

```{r}

unique(
  df_Bangladesh$icon
)
```

# 1. Defining the problem:

> In the weather data-set of Bangladesh there are 112 observations. The observations are daily based. We have found that there are 3 states during this time period, they are: a day can be a "clear-day" or "partly-cloudy-day" and there could be "rain" on that day. So we can predict what will be weather forecast for next 7 days after a state being chosen.

### Finding the probability of the states from the data-set that we can build a transition matrix:

> Printing the total occurrences of each state.

```{r}
table(df_Bangladesh$icon)

clear_day = count(filter(df_Bangladesh,icon=="clear-day"))
clear_day

partly_Cloudy_day = count(filter(df_Bangladesh,icon=="partly-cloudy-day"))
partly_Cloudy_day

rainy_day = count(filter(df_Bangladesh,icon=="rain"))
rainy_day
```

## Finding the probability:

> We will choose the weather column in order to find the probability of each states.

```{r}
weather_type = df_Bangladesh$icon
weather_type
```

## Defining the transitions:

> As there are 3 states, there will be states \* states =3\*3 = 9 transition states and from each state there will be 3 states outgoing.

### If the weather is in state "Rain"

```{r}
rain_to_rain = 0
rain_to_clear = 0
rain_to_partly_cloudy = 0
```

### If the weather is in state "clear-day"

```{r}
clear_to_clear = 0
clear_to_rain = 0
clear_to_partly_cloudy = 0
```

### If the weather is in state "partly-cloudy-day"

```{r}
partly_cloudy_to_partly_cloudy = 0
partly_cloudy_to_rain = 0
partly_cloudy_clear = 0
```

### Counting the outgoing occurrences from each state:

> This loop will find the count from each state to other states in order to calculate the probability. We will run the loop till length-1 time in order to avoid array index out of bound error and to make the count accurate.

```{r}
len = length(weather_type)-1

for (i in 1:len){
  if (weather_type[i] == "rain"){
    if(weather_type[i +1] == "rain"){
      rain_to_rain = rain_to_rain + 1
    }
    else if(weather_type[i +1] == "clear-day"){
      rain_to_clear= rain_to_clear + 1
    }
    else{
      rain_to_partly_cloudy = rain_to_partly_cloudy + 1
    }
  }
  
  else if (weather_type[i] == "clear-day"){
    if(weather_type[i +1] == "clear-day"){
      clear_to_clear = clear_to_clear + 1
    }
    else if(weather_type[i +1] == "rain"){
      clear_to_rain= clear_to_rain + 1
    }
    else{
      clear_to_partly_cloudy = clear_to_partly_cloudy + 1
    }
  }
  
  else{
    if(weather_type[i +1] == "partly-cloudy-day"){
      partly_cloudy_to_partly_cloudy = partly_cloudy_to_partly_cloudy + 1
    }
    else if(weather_type[i +1] == "rain"){
      partly_cloudy_to_rain= partly_cloudy_to_rain + 1
    }
    else{
      partly_cloudy_clear = partly_cloudy_clear + 1
    }
  }

}
```

### Printing the state transition counts from each state:

### If the weather is in state "Rain":

```{r}
rain_to_rain
rain_to_clear
rain_to_partly_cloudy
```

### If the weather is in state "clear-day":

```{r}
clear_to_clear
clear_to_rain
clear_to_partly_cloudy
```

### If the weather is in state "partly-cloudy-day":

```{r}
partly_cloudy_to_partly_cloudy
partly_cloudy_to_rain
partly_cloudy_clear
```

### Correcting the total count:

> Earlier we have calculated how many observations are there for each state. But from one state there can be no outgoing state to more than one states. So we have to calculate to transition from each state.

```{r}
rainy_transition = rain_to_rain + rain_to_clear + rain_to_partly_cloudy
clear_transition = clear_to_clear + clear_to_partly_cloudy + clear_to_rain
partly_transition = partly_cloudy_to_partly_cloudy + partly_cloudy_clear + partly_cloudy_to_rain
```

## Defining the probability for transition matrix:

### Transition probability from rain to other states:

```{r}
p_rain_to_rain = rain_to_rain/rainy_transition
p_rain_to_clear = rain_to_clear/rainy_transition
p_rain_to_partly_cloudy = rain_to_partly_cloudy/rainy_transition
```

> We can see the transition probabilities:

```{r}
p_rain_to_rain
p_rain_to_clear
p_rain_to_partly_cloudy
```

> We can also see the total probability of each row is 1 or not

```{r}
p_rain_to_clear + p_rain_to_rain + p_rain_to_partly_cloudy
```

### Transition probability from clear day to other states:

```{r}
p_clear_to_clear = clear_to_clear/clear_transition
p_clear_to_rain = clear_to_rain/clear_transition
p_clear_to_partly_cloudy = clear_to_partly_cloudy/clear_transition
```

> We can see the transition probabilities:

```{r}
p_clear_to_clear
p_clear_to_rain
p_clear_to_partly_cloudy
```

> We can also see the total probability of each row is 1 or not

```{r}
p_clear_to_clear + p_clear_to_rain + p_clear_to_partly_cloudy
```

### Transition probability from partly-cloudy-day to other states:

```{r}
p_partly_cloudy_to_partly_cloudy = partly_cloudy_to_partly_cloudy/partly_transition
p_partly_cloudy_to_rain = partly_cloudy_to_rain/partly_transition
p_partly_cloudy_clear = partly_cloudy_clear/partly_transition
```

> We can see the transition probabilities:

```{r}
p_partly_cloudy_to_partly_cloudy
p_partly_cloudy_to_rain
p_partly_cloudy_clear
```

> We can also see the total probability of each row is 1 or not

```{r}
p_partly_cloudy_to_partly_cloudy + p_partly_cloudy_clear + p_partly_cloudy_to_rain

```

# 2. Selecting the Markov Chain Model:

> I have imported the "markovchain" library to analyse the weather data.
>
> Now applying markov chain

### Building the transition matrix for markov chain model

> Transition matrix is essential to make a markov chain model. I calculated 1-step transition probabilities from each state manually. Now I will create a matrix using those probabilities. In this matrix byrow = True is enabled that will make the matrix created by row values.

```{r}
transition_matrix_weather <- round(matrix(c(p_clear_to_clear,p_clear_to_rain,p_clear_to_partly_cloudy,
                p_rain_to_clear,p_rain_to_rain, p_rain_to_partly_cloudy,
                p_partly_cloudy_clear, p_partly_cloudy_to_rain,
                p_partly_cloudy_to_partly_cloudy),nrow = 3, byrow = TRUE),3)
```

> We can see the transition matrix now.

```{r}
stateNames <- c("Clear Day","Rainy","Partly Cloudy Day")
row.names(transition_matrix_weather) <- stateNames; colnames(transition_matrix_weather) <- stateNames

transition_matrix_weather
```

### Making the Discrete Markov chain model:

> Using a markovchain object, here I have built a discrete markov chain model.

```{r}
discrete_transition_matrix_weather <- new("markovchain",transitionMatrix=transition_matrix_weather, 
             states=c("Clear Day","Rainy","Partly Cloudy Day"),
             name="MarkovChain for Weather Prediction") 
```

> We can see the markov chain model now.

```{r}
discrete_transition_matrix_weather
```

> We can also see the dimension of the matrix.

```{r}
dim(discrete_transition_matrix_weather)
```

### Plotting the transition matrix

```{r}
plotmat(t(transition_matrix_weather),pos = c(2,1), 
        lwd = 1, box.lwd = 2, 
        cex.txt = 0.8, 
        box.size = 0.12, 
        box.type = "circle", 
        box.prop = 0.7,
        box.col = "green",
        arr.length=.4,
        arr.width=.2,
        self.cex = .6,
        self.shifty = -.01,
        self.shiftx = .17,
        main = "Forecasting Rain on the Next Day")
```

# Forecasting of rain for next 7 days

### Defining Current State

> Let's take clear day as our current state , that is the probability that the weather goes from Clear day to Rainy day in the next 7 days I defined my state row matrix as this order **`: "Clear Day","Rainy","Partly Cloudy Day"`**, so the vector for the vector for choosing a clear Day is: **`(1,0,0)`**

```{r}
initial_state <- matrix(c(1,0,0),nrow=1, byrow=TRUE)
```

### Forecasting Rain in the next day:

```{r}
days <- 1
after_one_day <- round(initial_state * discrete_transition_matrix_weather ^ days, 3)
after_one_day

after_one_day <- round(transition_matrix_weather%^%days,3)
after_one_day
#plot
plotmat(t(after_one_day),pos = c(2,1), 
        lwd = 1, box.lwd = 2, 
        cex.txt = 0.8, 
        box.size = 0.12, 
        box.type = "circle", 
        box.prop = 0.7,
        box.col = "green",
        arr.length=.4,
        arr.width=.2,
        self.cex = .6,
        self.shifty = -.01,
        self.shiftx = .17,
        main = "Forecasting Rain on the Next Day")
```

> Here we can see that there is only 5.7% chance of raining in the next day of a clear day.

### Forecasting Rain after 2 days:

```{r}
days <- 2
after_two_days <- round(initial_state * discrete_transition_matrix_weather ^ days,3)
after_two_days

after_two_days <- round(transition_matrix_weather%^%days,3)
after_two_days

plotmat(t(after_two_days),pos = c(1,2), 
        lwd = 1, box.lwd = 2, 
        cex.txt = 0.8, 
        box.size = 0.1, 
        box.type = "circle", 
        box.prop = 0.5,
        box.col = "green",
        arr.length=.1,
        arr.width=.1,
        self.cex = .6,
        self.shifty = -.01,
        self.shiftx = .14,
        main = "Forecasting Rain After 2 Days")
```

> Here we can see that there is only 10.5% chance of raining after 2 days of a clear day.

### Forecasting Rain after 3 days

```{r}
days<-3
after_three_days <- round(initial_state * discrete_transition_matrix_weather ^ days,3)
after_three_days

after_three_days <- round(transition_matrix_weather%^%days,3)
after_three_days

plotmat(t(after_three_days),pos = c(1,2), 
        lwd = 1, box.lwd = 2, 
        cex.txt = 0.8, 
        box.size = 0.1, 
        box.type = "circle", 
        box.prop = 0.5,
        box.col = "green",
        arr.length=.1,
        arr.width=.1,
        self.cex = .6,
        self.shifty = -.01,
        self.shiftx = .14,
        main = "Forecasting Rain After 3 Days")
```

> Here we can see that there is only 13.7% chance of raining after 3 days of a clear day.

### Forecasting Rain after 4 days

```{r}
days <- 4
after_four_days <- round(initial_state * discrete_transition_matrix_weather ^ days,3)
after_four_days

after_four_days <- round(transition_matrix_weather%^%days,3)
after_four_days

plotmat(t(after_four_days),pos = c(1,2), 
        lwd = 1, box.lwd = 2, 
        cex.txt = 0.8, 
        box.size = 0.1, 
        box.type = "circle", 
        box.prop = 0.5,
        box.col = "green",
        arr.length=.1,
        arr.width=.1,
        self.cex = .6,
        self.shifty = -.01,
        self.shiftx = .14,
        main = "Forecasting Rain After 4 Days")

```

> Here we can see that there is only 15.7% chance of raining after 4 days of a clear day.

### Forecasting Rain after 5 days

```{r}
days <- 5
after_five_days <- round(initial_state * discrete_transition_matrix_weather ^ days,3)

after_five_days

after_five_days <- round(transition_matrix_weather%^%days,3)
after_five_days

plotmat(t(after_five_days),pos = c(1,2), 
        lwd = 1, box.lwd = 2, 
        cex.txt = 0.8, 
        box.size = 0.1, 
        box.type = "circle", 
        box.prop = 0.5,
        box.col = "green",
        arr.length=.1,
        arr.width=.1,
        self.cex = .6,
        self.shifty = -.01,
        self.shiftx = .14,
        main = "Forecasting Rain After 5 Days")
```

> Here we can see that there is only 16.7% chance of raining after 5 days from a clear day.

### Forecasting Rain after 6 days

```{r}
days <- 6
after_six_days <- round(initial_state * discrete_transition_matrix_weather ^ days,3)

after_six_days

after_six_days <- round(transition_matrix_weather%^%days,3)
after_six_days

plotmat(t(after_six_days),pos = c(1,2), 
        lwd = 1, box.lwd = 2, 
        cex.txt = 0.8, 
        box.size = 0.1, 
        box.type = "circle", 
        box.prop = 0.5,
        box.col = "green",
        arr.length=.1,
        arr.width=.1,
        self.cex = .6,
        self.shifty = -.01,
        self.shiftx = .14,
        main = "Forecasting Rain After 6 Days")
```

> Here we can see that there is only 17.2% chance of raining after 6 days from a clear day.

### Forecasting Rain after 7 days

```{r}

days <- 7 
after_seven_days <- round(initial_state * discrete_transition_matrix_weather ^ days,3)

after_seven_days

after_seven_days <- round(transition_matrix_weather%^%days,3) 
after_seven_days

plotmat(t(after_seven_days),pos = c(1,2),
        lwd = 1, box.lwd = 2, cex.txt = 0.8,
        box.size = 0.1, box.type = "circle",
        box.prop = 0.5,
        box.col ="green",
        arr.length=.1,
        arr.width=.1,
        self.cex = .6,
        self.shifty = -.01,
        self.shiftx = .14,
        main = "Forecasting Rain After 7 Days")
```

> Here we can see that there is only 17.5% chance of raining after 7 days from a clear day.

## Simulation of Multi Nomial Markov Chain on Weather Data:

We will use transition matrix transition_matrix_weather to simulate.

```{r}
round(transition_matrix_weather,2)
```

```{r}
run.mc.sim <- function( P, t) { 

  # number of possible states

  num.states <- nrow(P)
  print("num.states")
  print(num.states)
  # stores the states X_t through time

  states <- numeric(t)
  print('the value of state is:')
  print(states)
  # initialize variable for first state 

  states[1]    <- 1


  for(i in 2:t) { #i=2

    # probability vector to simulate next state X_{t+1}

    a<-states[i-1]
    print("the value of a is")
    print( a)
    pt  <- P[a, ]
    print("THe value of pt")
    print(pt)
    ## draw from multinomial and determine state

    states[i] <-  which(rmultinom(1, 1, pt) == 1)
    print("value of states after rmul")
    print(states[i])
  }

  return(states)

}
```

```{r}

# setup transition matrix p, here we will use transition_matrix_weather

P <- transition_matrix_weather

#num.iterations 

t=100

MC<-run.mc.sim(P,t)

MC
typeof(MC)
length(MC)
plot(MC,type="b",col='green')
```

# Problem 2: Markov chain on Stock price data

```{python}
import pandas as pd

```

```{python}
# this is a stock data set of dhaka stcok exchange from 2021
df_2 = pd.read_json("C:\\Users\\USER\\Downloads\\STA301\\prices_2021.json")
#will print unique company names in the dataset
df_2["trading_code"].unique() 
df_2.head(10)#viewing fisrt 10 rows
df_2.columns # will show the column/feature names of the stock dataset.
#with pd.option_context("display.max_columns=12", None):
   # display(df_2.head(10))
```

```{python}
BATBC = df_2.loc[df_2['trading_code']=='BATBC']

BATBC['closing_price']
BATBC = BATBC[['date','trading_code','closing_price']]
BATBC['date'] = pd.to_datetime(BATBC['date']).dt.date
BATBC.head(20)
BATBC.to_csv('BATBC.csv', index = True)
```

# Analysis of Stock Market Trend Analysis of British American Tobacco Bangladesh

**Importing the data-set:**

```{r}
df_batbc = read.csv("C:\\Users\\USER\\Downloads\\STA301\\BATBC.csv")
head(df_batbc,20)
```

```{r}
unique(
  df_batbc$closing_price
)
```

# 1. Abstract

Stock Market of Bangladesh is very much inconsistent in terms of closing rates and returning rates. This paper aims to analyze the stock price trend of British American Tobacco Bangladesh Company(BATBC). We will use Markov Chain model to analyze this trend.

# 2. Data Collection:

This stock data set covers the period from January 2021 to December 2021 on the Dhaka Stock Exchange. Then, we limited our observations to those made where BATBC is the Company and made a separate data-set containing only the observations of BATBC.

# 3. Results:

The BATBC data set has 240 observations. The stock's daily closing price will be taken into account. Among these 240 observations, 221 stock closing prices are distinct. We will investigate three different scenarios in order to examine the trend: if the stock closing price is higher or lower the day after a specific stock closing price, or if it stays the same.

## 3.1 Determining the states:

There will be three states in total in the transition matrix. If the stock closing price is higher than a specific day, we will denote higher, if it is lower then lower, otherwise if it remains same then constant.

## 3.2 Finding the transitional probability:

We encoded the closing_price column into those three states. We took first observation as constant_closing_price and then if a stock closing price was higher than current one, then we encoded as higher_closing_price, and if it decreased, then it was encoded as lower_closing_price.

```{r}
print("hello world")
closing_price_encoded <- c("constant")
closing_price_encoded

closing_price = df_batbc$closing_price
len_closing_price = length(df_batbc$closing_price)-1

len_closing_price

for (i in 1:len_closing_price){
  if (closing_price[i] == closing_price[i+1]){
    closing_price_encoded <- c(closing_price_encoded,"constant")
  }
  
  else if (closing_price[i] < closing_price[i+1]){
    closing_price_encoded <- c(closing_price_encoded,"Higher")
  }
  
  
  else{
    closing_price_encoded <- c(closing_price_encoded,"Lower")
  }

}
closing_price_encoded
typeof(closing_price_encoded)
length(closing_price_encoded)
```

Now replacing the closing_price column with encoded closing price array

```{r}
df_batbc$closing_price <- closing_price_encoded

```

## **Defining the transitions:**

> As there are 3 states, there will be states \* states =3\*3 = 9 transition states and from each state there will be 3 states outgoing.

**If the stock closing price is in state "Constant":**

```{r}
constant_to_constant = 0
constant_to_higher = 0
constant_to_lower = 0
```

**If the stock closing price is in state "Lower":**

```{r}
lower_to_constant = 0
lower_to_higher = 0
lower_to_lower = 0
```

**If the stock closing price is in state "Higher":**

```{r}
higher_to_constant = 0
higher_to_higher = 0
higher_to_lower = 0
```

### **Counting the outgoing occurrences from each state:**

> This loop will find the count from each state to other states in order to calculate the probability. We will run the loop till length-1 time in order to avoid array index out of bound error and to make the count accurate.

```{r}

closing_price = df_batbc$closing_price
len_closing_price = length(df_batbc$closing_price)-1
for (i in 1:len_closing_price){
  if (closing_price[i] == "constant"){
    if(closing_price[i +1] == "constant"){
      constant_to_constant = constant_to_constant + 1
    }
    else if(closing_price[i +1] == "Lower"){
      constant_to_lower = constant_to_lower + 1
    }
    else{
      constant_to_higher = constant_to_higher + 1
    }
  }
  
  else if (closing_price[i] == "Lower"){
    if(closing_price[i +1] == "Lower"){
      lower_to_lower = lower_to_lower + 1
    }
    else if(closing_price[i +1] == "constant"){
      lower_to_constant = lower_to_constant + 1
    }
    else{
      lower_to_higher = lower_to_higher + 1
    }
  }
  
  else{
    if(closing_price[i +1] == "Higher"){
      higher_to_higher = higher_to_higher + 1
    }
    else if(closing_price[i +1] == "Lower"){
      higher_to_lower = higher_to_lower + 1
    }
    else{
      higher_to_constant = higher_to_constant + 1
    }
  }

}
```

```{r}
constant_to_constant
constant_to_higher 
constant_to_lower

lower_to_constant
lower_to_higher
lower_to_lower

higher_to_constant
higher_to_higher
higher_to_lower

```

### **Correcting the total count:**

> Earlier we have calculated how many observations are there for each state. But from one state there can be no outgoing state to more than one states. So we have to calculate to transition from each state.

```{r}
constant_transition = constant_to_constant + constant_to_higher + constant_to_lower

lower_transition = lower_to_constant + lower_to_higher + lower_to_lower

higher_transition = higher_to_constant + higher_to_higher + higher_to_lower

constant_transition 
lower_transition
higher_transition
```

**Defining the probability for transition matrix:**

### **Transition probability from constant to other states:**

```{r}
p_constant_to_constant = constant_to_constant/ constant_transition
p_constant_to_higher = constant_to_higher/ constant_transition
p_constant_to_lower = constant_to_lower/ constant_transition

p_constant_to_constant
p_constant_to_higher 
p_constant_to_lower
```

### **Transition probability from Lower to other states:**

```{r}
p_lower_to_constant = lower_to_constant / lower_transition
p_lower_to_higher = lower_to_higher / lower_transition
p_lower_to_lower = lower_to_lower / lower_transition

p_lower_to_constant
p_lower_to_higher
p_lower_to_lower
```

### **Transition probability from Higher to other states:**

```{r}
p_higher_to_constant = higher_to_constant / higher_transition
p_higher_to_higher = higher_to_higher / higher_transition
p_higher_to_lower = higher_to_lower / higher_transition

p_higher_to_constant
p_higher_to_higher
p_higher_to_lower
```

### 

# **2. Selecting the Markov Chain Model:**

> I have imported the "markovchain" library to analyse the weather data.
>
> Now applying markov chain

### **Building the transition matrix for markov chain model**

> Transition matrix is essential to make a markov chain model. I calculated 1-step transition probabilities from each state manually. Now I will create a matrix using those probabilities. In this matrix byrow = True is enabled that will make the matrix created by row values.

```{r}
transition_matrix_stock <- matrix(c(p_higher_to_higher, p_higher_to_lower,p_higher_to_constant, p_lower_to_higher , p_lower_to_lower,p_lower_to_constant , p_constant_to_higher , p_constant_to_lower,p_constant_to_constant),nrow = 3, byrow = TRUE)

```

We can see the transition matrix now.

```{r}
stateNames_stock <- c("Higher","Lower","Constant")
row.names(transition_matrix_stock) <- stateNames_stock; colnames(transition_matrix_stock) <- stateNames_stock

round(transition_matrix_stock,3)
```

### **Making the Discrete Markov chain model:**

> Using a Markov chain object, here I have built a discrete markov chain model.

```{r}
discrete_transition_matrix_stock <- new("markovchain",transitionMatrix=transition_matrix_stock, 
             states=c("Higher","Lower","Constant"),
             name="MarkovChain for Stock Price trend Analysis")
discrete_transition_matrix_stock
```

### **Plotting the transition matrix**

```{r}
plotmat(t(round(transition_matrix_stock,2)),pos = c(2,1), 
        lwd = 1, box.lwd = 2, 
        cex.txt = 0.8, 
        box.size = 0.12, 
        box.type = "circle", 
        box.prop = 0.7,
        box.col = "green",
        arr.length=.4,
        arr.width=.2,
        self.cex = .6,
        self.shifty = -.01,
        self.shiftx = .17,
        main = "One step transition probability")
```

## The code chunk below is built in Markov chain transition matrix and model builder:

I have added this code to cross check my own calculated transition matrix with the built in library's result. Here we can see the results are same.

```{r}

sequence <- closing_price

mcFit <- markovchainFit(data=sequence)
mcFit

initial_state <- matrix(c(1,0,0),nrow=1, byrow=TRUE)

```

# **Analysis of stock closing price trend after next 7 days**

### **Defining Current State**

> Let's take clear day as our current state , that is the probability that the stock closing price goes from Lower to High after the next 7 days. I defined my state row matrix as this order **`: "Higher","Lower","Constant",`**, so the vector for the vector for choosing a clear Day is: **`(0,1,0)`**

```{r}
initial_state_stock <- matrix(c(0,1,0),nrow=1, byrow=TRUE)
```

## **Analysis of stock closing price trend after 2 days**

```{r}
days <- 2
after_two_days_stock <- round(initial_state_stock * discrete_transition_matrix_stock ^ days,2)
after_two_days_stock
```

```{r}
after_two_days_stock <- round(transition_matrix_stock%^%days,2)
after_two_days_stock
```

```{r}
plotmat(t(round(after_two_days_stock,2)),pos = c(1,2), 
        lwd = 1, box.lwd = 2, 
        cex.txt = 0.8, 
        box.size = 0.1, 
        box.type = "circle", 
        box.prop = 0.5,
        box.col = "green",
        arr.length=.1,
        arr.width=.1,
        self.cex = .6,
        self.shifty = -.01,
        self.shiftx = .14,
        main = "Analysis of stock closing price trend after 2 days")
```

## **Analysis of stock closing price trend after 7 days**

```{r}
days <- 7
after_seven_days_stock <- round(initial_state_stock * discrete_transition_matrix_stock ^ days,2)
after_seven_days_stock
```

```{r}
after_seven_days_stock <- round(transition_matrix_stock%^%days,2)
after_seven_days_stock
```

```{r}
plotmat(t(round(after_seven_days_stock,2)),pos = c(1,2), 
        lwd = 1, box.lwd = 2, 
        cex.txt = 0.8, 
        box.size = 0.1, 
        box.type = "circle", 
        box.prop = 0.5,
        box.col = "green",
        arr.length=.1,
        arr.width=.1,
        self.cex = .6,
        self.shifty = -.01,
        self.shiftx = .14,
        main = "Analysis of stock closing price trend after 7 days")
```

## Simulation of Multi Nomial Markov Chain on BATBC Stock Price:

> We will use transition matrix `transition_matrix_stock` to simulate

```{r}
round(transition_matrix_stock,2)
```

```{r}
simulation_for_stock <- function( P, t) { 

  # number of possible states

  num.states <- nrow(P)
  print("num.states")
  print(num.states)
  # stores the states X_t through time

  states <- numeric(t)
  print('the value of state is:')
  print(states)
  # initialize variable for first state 

  states[1]    <- 1


  for(i in 2:t) { #i=2

    # probability vector to simulate next state X_{t+1}

    a<-states[i-1]
   
    pt  <- P[a, ]
   
    ## draw from multinomial and determine state

    states[i] <-  which(rmultinom(1, 1, pt) == 1)
   
  }

  return(states)

}
```

```{r}
# setup transition matrix p, here we will use transition_matrix_weather

P <- transition_matrix_stock

#num.iterations 

t=100

simulation_for_stock_states<-simulation_for_stock(P,t)
print("simulation_for_stock_states:")
simulation_for_stock_states

plot(simulation_for_stock_states,type="b",col='green')
```
